Project Uccle

### 2020-05-22 project.Rmd
more analysis and better Rmd output

```{r libraries, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F, fig.height = 360)
library(stringi) # %s+%
library(lubridate) # floor_date
library(xts) # zoo
library(fpp) # auto.arima
# library(gridExtra) # for gg-plotting multiple plots
library(cowplot) # for plot_grid
library(tidyverse) # map_df, read_csv, group_by, summarize
library(tsbox) # https://cran.r-project.org/web/packages/tsbox/vignettes/tsbox.html
print_dec <- function(x, decimals=2) trimws(format(round(x, decimals), nsmall=decimals))

@@

```{r import dataset, include=FALSE}
# MEMS
domain <- c(-7200,7200); pattern <- "MEM.*.csv"; name <- "MEM"; description <- "Membach"
# UCCS
domain <- c(-350000,350000); pattern <- "UCCS.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"
seis <-
  list.files(path="csv/", pattern=pattern) %>%
  map_df(~read_csv("csv/" %s+% .))
```

@@

block_start <- NULL
block_end <- NULL
blocks <- data.frame(start=integer(), end=integer())
for (i in 1:length(seis.missing.ids)) {
  if (is.null(block_start)) {
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  } else if ((block_end + 1) == seis.missing.ids[i]) {
    block_end <- seis.missing.ids[i]
  } else {
    blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  }
}
blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
blocks$interval <- hours(blocks$end - blocks$start + 1)
blocks$start <- seis.first.second + hours(blocks$start)
blocks$end <- seis.first.second + hours(blocks$end)
plot(c(blocks$start, blocks$end), rep(hour(blocks$interval), 2),
     col=ifelse(blocks$interval<3600, "darkgreen", "darkred"),
     main="Consecutive hours of missing values",
     ylab="Hours", xlab="Interval start",
     type="l", lwd=10,
     xaxt="n", xlim=c(min(seis$utc_second), max(seis$utc_second))
     )
axis.POSIXct(1, at=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = "1 month"), format="%m-%y", cex=0.8)
text(x=blocks$start, y=hour(blocks$interval), labels = paste0(format(blocks$start, "%y-%m-%d %H:%M")), cex=0.9, pos=2)
text(x=blocks$end, y=hour(blocks$interval), labels = paste0(format(blocks$end, "%y-%m-%d %H:%M")), cex=0.9, pos=4)
text(x=blocks$start + blocks$interval/2, y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=3)
remove(seis.complete, seis.missing.ids)

Box.test(forec$model$residuals, type="Box-Pierce")

@@

```{r plot data with (p)acf}
seis.h.start <- as_datetime('2018-10-29T00:00:00')
seis.h.end <- as_datetime('2020-03-01T00:00:00')
seasonality.period <- 24
seasonality.xlab <- "Days"
seasonality.aggregate <- function(ds) na.aggregate(seis.h.ts, function(x) floor(x/7))
seasonality.period <- 168
seasonality.xlab <- "Weeks"
seasonality.aggregate <- function(ds) na.aggregate(ds, floor)
seis.h.ts <- build_ts(seis.h, periods = c(seasonality.period), start = seis.h.start, end = seis.h.end, force_ms = T)
plot_grid(
  autoplot(seis.h.ts, main = "Hourly seismic movement", xlab = seasonality.xlab, ylab = "mean movement (nm)"),
  gghistogram(seis.h.ts) + xlab("mean movement (nm)"),
  nrow = 2)
plot_grid(
  autoplot(acf(seis.h.ts, na.action=na.pass, plot=F, lag.max=seasonality.period)) + scale_x_continuous(),
  autoplot(pacf(seis.h.ts, na.action=na.pass, plot=F, lag.max=seasonality.period)) + scale_x_continuous(),
  nrow = 2)

@@

  mse <- mean(diff^2)
  mae <- mean(abs(diff))
  rmse <- sqrt(mse)
  R2 <- 1-(sum((diff)^2)/sum((base-mean(base))^2))
  errorstring <- paste0("MSE=", mse, "    MAE=", mae, "    RMSE=", rmse, "    R2=", R2)
  print(errorstring)

@@

plot_grid(
  autoplot(aggregated) + ylab("") + xlab(""),
  autoplot(ts(season.24$seasonal, frequency=24)) + ylab("") + xlab(""),
  autoplot(ts(season.168$seasonal, frequency=24)) + ylab("") + xlab(""),
  autoplot(ts(season.168$trend, frequency=24)) + ylab("") + xlab(""),
  autoplot(aggregated-season.24$seasonal-season.168$seasonal-season.168$trend) + ylab("") + xlab(""),
  nrow = 5)

@@

plot_grid(
  autoplot(aggregated) + ylab("") + xlab(""),
  autoplot(ts(season.24$seasonal, frequency=24)) + ylab("") + xlab(""),
  autoplot(ts(season.168$seasonal, frequency=24)) + ylab("") + xlab(""),
  autoplot(ts(season.168$trend, frequency=24)) + ylab("") + xlab(""),
  autoplot(aggregated-season.24$seasonal-season.168$seasonal-season.168$trend) + ylab("") + xlab(""),
  nrow = 5)

@@

plot_grid(
  autoplot(ts(seasonbase.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
      geom_line(aes(y=season.24$figure, col="red"), show.legend = F),
  autoplot(ts(seasonbase.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() +
      geom_line(aes(y=season.168$figure, col="red")) +
      geom_line(aes(y=season.168$figure + rep(season.24$figure, 7), col="blue")),
  nrow = 2)

@@

```{r test_ts}
seis.test.start <- as_datetime('2019-05-20T00:00:00')
seis.test.end <- as_datetime('2020-03-02T00:00:00')

seis.h.test <- build_ts(seis.h, periods = c(seasonality.period), start = seis.test.start, end = seis.test.end, force_ms = T)
```

```{r diffs acf/pacf}
seis.h.test %>% ggtsdisplay(lag.max = 168)
seis.h.test %>% diff() %>% ggtsdisplay(lag.max = 168*5)
# seis.h.test %>% diff(D=2) %>% ggtsdisplay(lag.max = 168)
# seis.h.test %>% diff(lag=24) %>% ggtsdisplay(lag.max = 168)
# seis.h.test %>% diff() %>% diff(lag=24) %>% ggtsdisplay(lag.max = 24, main="Supposed model: ARIMA(24,1,0)(0,1,2)[24]") # ARIMA(0,1,0)(0,1,2)[24]
# previous => ARIMA(5,1,0)(0,1,7)[24]
# seis.h.test %>% diff() %>% diff(lag=24, D=2) %>% ggtsdisplay(lag.max = 168)
seis.h.test %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5) # arima (0,0,0)(0,1,1/2)[168]
seis.h.test %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5) # arima (0,1,0)(0,1,1)[168]
# seis.h.test %>% diff() %>% diff(lag=168, D=2) %>% ggtsdisplay(lag.max = 168)
seis.h.test %>% diff() %>% diff(lag=24) %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168)
```

```{r arima}

plot_forecast <- function(forec) {
  BP <- Box.test(forec$model$residuals, type="Box-Pierce")
  autoplot(forec) +
    labs(title=forec$method, subtitle = paste0("Box-Pierce p-value=", print_dec(BP$p.value, decimals=4))) +
    autolayer(forec$mean, series="Forecast") -> plotted
  return(plotted)
}

to_test <- list(
  #c(0,0,0,0,0,1),  # BAD, Box-Pierce = 0
  #c(0,0,0,0,0,2),  # BAD, Box-Pierce = 0
  #c(0,0,0,0,1,1),  # nice, Box-Pierce = 0
  c(0,0,0,0,1,2),
  #c(0,0,0,0,1,3),  # CAN'T BE DONE: 3*168 = 504 > 350 (limit of arima on r)
  #c(0,1,0,0,0,1),  # BAD, Box-Pierce = 0
  c(0,1,0,0,1,1),
  c(0,1,0,0,1,2),
  c(2,0,0,0,1,0),  # hint from auto.arima D=1
  c(5,0,0,0,1,0),  # hint from auto.arima D=1
  c(3,1,1,0,1,0),  # hint from auto.arima D=1 d=1
  c(3,1,1,0,1,1),  # after hint then ACF + PACF
)
if (!exists("models")){
  models <- list()
}
for (model in to_test) {
  model <- c(3,1,1,0,1,1)
  fit.name <- paste0("fit", paste0(model, collapse = ""))
  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.test) %s+% "]")
  models[[fit.name]] <- Arima(seis.h.test, model[1:3], model[4:6])
  models[[forecast.name]] <- forecast(models[[fit.name]], h=168*10)
}
models[["fitauto"]] <- auto.arima(seis.h.test, D=1, approx=F, trace = T)
models[["forecastauto"]] <- forecast(models[["fitauto"]], h=168*10)
models[["fit311010"]] <- models[["fitauto"]]
models[["forecast311010"]] <- forecast(models[["fit311010"]], h=168*10)

@@

checkresiduals(plotting.model)
residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
plot_grid(residuals.acf, residuals.pacf, nrow=2)
plot_forecast(plotting.forecast)
plots <- show_overlap_and_diff(
  seis.h.test, plotting.forecast$fitted, xlab=seasonality.xlab,
  base.desc = "original series",
  compare.desc = "reconstructed series",
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="% over ts span",
  diff.ylab.sec.accuracy = .01
)
plot_grid(plots[[1]], plots[[2]], nrow = 2)

plotting.model <- models[["fit010012"]]
plotting.forecast <- models[["forecast010012"]]
checkresiduals(plotting.model)
residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
plot_grid(residuals.acf, residuals.pacf, nrow=2)
plot_forecast(plotting.forecast)
plots <- show_overlap_and_diff(
  seis.h.test, plotting.forecast$fitted, xlab=seasonality.xlab,
  base.desc = "original series",
  compare.desc = "reconstructed series",
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="% over ts span",
  diff.ylab.sec.accuracy = .01
)
plot_grid(plots[[1]], plots[[2]], nrow = 2)

```{r}
(seis.test.start - seis.first.second)

deseasoned24 <- seis.h.ts - seasonality
autoplot(deseasoned24)
frequency(deseasoned24)
de24.test <- msts(window(deseasoned24, start=205), seasonal.periods = c(168))
frequency(de24.test)
autoplot(de24.test)
de24.test %>% ggtsdisplay(lag.max = 168*2)
de24.test %>% diff() %>% ggtsdisplay(lag.max = 168*2)
de24.test %>% diff(D=2) %>% ggtsdisplay(lag.max = 168*2)
de24.test %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*2)
de24.test %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*2, main="Supposed model: ARIMA(0,1,0)(0,1,1)[168]")
de24.test %>% diff() %>% diff(lag=168, D=2) %>% ggtsdisplay(lag.max = 168*2)

if (!exists("fit010011de")){
  fit010011de <- Arima(de24.test, c(0,1,0), c(0,1,1))
}
checkresiduals(fit010011de)
residuals.acf <- autoplot(acf(fit010011de$residuals, na.action = na.pass, plot=F, lag.max=168*3)) + scale_x_continuous()
residuals.pacf <- autoplot(pacf(fit010011de$residuals, na.action = na.pass, plot=F, lag.max=168*3)) + scale_x_continuous()
plot_grid(residuals.acf, residuals.pacf, nrow=2)
if (!exists("fit010011de.forec")){
  fit010011de.forec <- forecast(fit010011de, h=168*10)
}
plot_forecast(forecast(fit010011de.forec, h=168*10))

# NO  fit011011de <- Arima(de24.test, c(0,1,1), c(0,1,1))
if (!exists("fit011011de")){
  fit011011de <- Arima(de24.test, c(0,1,1), c(0,1,1))
  arima
}
fit010011de$aic
fit011011de$aic
checkresiduals(fit011011de)
residuals.acf <- autoplot(acf(fit011011de$residuals, na.action = na.pass, plot=F, lag.max=168*3)) + scale_x_continuous()
residuals.pacf <- autoplot(pacf(fit011011de$residuals, na.action = na.pass, plot=F, lag.max=168*3)) + scale_x_continuous()
plot_grid(residuals.acf, residuals.pacf, nrow=2)
if (!exists("fit011011de.forec")){
  fit011011de.forec <- forecast(fit011011de, h=168*10)
}
plot_forecast(forecast(fit011011de.forec, h=168*10))

```{r}

seis.test.start <- as_datetime('2019-05-20T00:00:00')
seis.test.end <- as_datetime('2020-03-02T00:00:00')

seis.h.test <- build_ts(seis.h, periods = c(seasonality.period), start = seis.test.start, end = seis.test.end, force_ms = T)

# First seasonal difference (quarterly)
# seasonality.arima <- 24
D<-1
seis.h.test.ds<-diff(seis.h.test,differences=D)
plot(seis.h.test.ds)
abline(h=mean(seis.h.test.ds), col="red", lty="dotted")
sd(seis.h.test.ds)
tsdisplay(seis.h.test.ds, lag.max = 168*2)
autoplot(acf(seis.h.test.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
autoplot(pacf(seis.h.test.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
seis.h.test.ds<-diff(seis.h.test,lag=seasonality.period,differences=D)
plot(seis.h.test.ds)
abline(h=mean(seis.h.test.ds), col="red", lty="dotted")
sd(seis.h.test.ds)
autoplot(acf(seis.h.test.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
autoplot(pacf(seis.h.test.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
tsdisplay(seis.h.test.ds, lag.max = 168*2)
# pacf cuts off after p=168
```

```{r}
fit <- auto.arima(seis.h.test, d=1, D=1, trace = T, parallel = T)
```
 Fitting models using approximations to speed things up...

@@

```{r}
library(car)

autoplot(acf(seis.h.test.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()
autoplot(pacf(seis.h.test.ds, plot=F, lag.max = 168*2)) + scale_x_continuous()

>>>>

```{r libraries, include=FALSE}
library(knitr)
opts_chunk$set(warning = F, message = F)
library(stringi) # %s+%
library(lubridate) # floor_date
library(xts) # zoo
library(fpp) # auto.arima
# library(gridExtra) # for gg-plotting multiple plots
library(cowplot) # for plot_grid
library(car) # qqPlot
library(tidyverse) # map_df, read_csv, group_by, summarize
library(tsbox) # https://cran.r-project.org/web/packages/tsbox/vignettes/tsbox.html
print_dec <- function(x, decimals=2) trimws(format(round(x, decimals), nsmall=decimals))

@@

```{r import dataset, include=FALSE}
if (!exists("domain")) {
  # UCCS
  domain <- c(-350000,350000); pattern <- "UCCS.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"; proj_file <- 'proj.uccs.RData'
  # MEMS
  domain <- c(-7200,7200); pattern <- "MEM.*\\.csv"; name <- "MEM"; description <- "Membach"; proj_file <- 'proj.mems.RData'
}
if (exists("proj_file")) {
  load(paste0("/media/data/Documents/Uni/spatiotemp/villalobos/project/", proj_file))
  # save.image(paste0("/media/data/Documents/Uni/spatiotemp/villalobos/project/", proj_file))
}
seis <-
  list.files(path="csv/", pattern=pattern) %>%
  map_df(~read_csv("csv/" %s+% .))
```

@@

block_start <- NULL
block_end <- NULL
blocks <- data.frame(start=integer(), end=integer())
for (i in 1:length(seis.missing.ids)) {
  if (is.null(block_start)) {
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  } else if ((block_end + 1) == seis.missing.ids[i]) {
    block_end <- seis.missing.ids[i]
  } else {
    blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  }
}
blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
blocks$interval <- hours(blocks$end - blocks$start + 1)
blocks$start <- seis.first.second + hours(blocks$start)
blocks$end <- seis.first.second + hours(blocks$end)
plot(c(rbind(blocks$start, blocks$end, NA)), rep(hour(blocks$interval), each=3),
     col=ifelse(blocks$interval<3600, "darkgreen", "darkred"),
     main="Consecutive hours of missing values",
     ylab="Hours", xlab="Interval start",
     type="l", lwd=10,
     xaxt="n", xlim=c(min(seis$utc_second), max(seis$utc_second))
     )
axis.POSIXct(1, at=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = "1 month"), format="%m-%y", cex=0.8)
text(x=blocks$start, y=hour(blocks$interval), labels = paste0(format(blocks$start, "%y-%m-%d %H:%M")), cex=0.9, pos=2)
text(x=blocks$end, y=hour(blocks$interval), labels = paste0(format(blocks$end, "%y-%m-%d %H:%M")), cex=0.9, pos=4)
text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=3)
remove(seis.complete, seis.missing.ids)

blocks$interval
Box.test(seis.h$mean_movement_s, type="Box-Pierce")

@@

```{r test_train split ts, include=FALSE}
seis.h.start <- as_datetime('2018-10-29T00:00:00')
seis.h.end <- as_datetime('2020-03-02T00:00:00') - seconds(1)
seasonality.period <- 24
seasonality.xlab <- "Days"
seasonality.aggregate <- function(ds) na.aggregate(seis.h.ts, function(x) floor(x/14))
seasonality.period <- 168
seasonality.xlab <- "Weeks"
seasonality.aggregate <- function(ds) na.aggregate(ds, floor)
seis.h.ts <- build_ts(seis.h, periods = c(seasonality.period), start = seis.h.start, end = seis.h.end, force_ms = T)

seis.test.split <- seis.h.start + days(52*7)

seis.train.start <- seis.h.start
seis.train.end <- seis.test.split - seconds(1)
seis.h.train <- build_ts(seis.h, periods = c(seasonality.period), start = seis.train.start, end = seis.train.end, force_ms = T)

seis.test.start <- seis.test.split
seis.test.end <- as_datetime('2020-05-04T00:00:00') - seconds(1)
seis.h.test <- build_ts(seis.h, periods = c(seasonality.period), start = seis.h.start, end = seis.test.end, force_ms = T)
seis.h.test <- window(seis.h.test, start = c(end(seis.h.train)[1] + 1, 1))
seis.test.length <- length(seis.h.test)
```

```{r plot data with (p)acf}
seis.h.24 <- build_ts(seis.h, periods = c(24), start = seis.h.start, end = seis.h.end)
seasonplot(seis.h.24, type="l", col=rainbow(7))
legend("topleft", legend=weekdays(seis.h.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)

plot_grid(
  autoplot(seis.h.ts, main = "Hourly seismic movement", xlab = seasonality.xlab, ylab = "mean movement (nm)"),
  gghistogram(seis.h.ts, add.normal = T) + xlab("mean movement (nm)"),
  nrow = 2)
plot_grid(
  autoplot(acf(seis.h.ts, na.action=na.pass, plot=F, lag.max=seasonality.period)) + scale_x_continuous(),
  autoplot(pacf(seis.h.ts, na.action=na.pass, plot=F, lag.max=seasonality.period)) + scale_x_continuous(),
  nrow = 2)

@@

  mse <- mean(diff^2, na.rm = T)
  mae <- mean(abs(diff), na.rm = T)
  rmse <- sqrt(mse)
  R2 <- 1-(sum((diff)^2, na.rm = T)/sum((base-mean(base))^2, na.rm = T))
  errorstring <- paste0("MSE=", mse, "    MAE=", mae, "    RMSE=", rmse, "    R2=", R2)
  print(errorstring)

@@

@@

# this third iteration isn't actually needed
season.24 <- trend_season(aggregated-season.168$seasonal, 24)
season.168 <- trend_season(aggregated-season.24$seasonal, 168)

plot_grid(
  autoplot(aggregated) + labs(title="Timeseries", cex=0.7) + ylab("") + xlab("") + theme(plot.margin=margin(b=-6, unit="pt")),
  autoplot(ts(season.24$figure, frequency=1)) + labs(title="Daily seasonality", cex=0.7) + ylab("") + xlab("") + theme(plot.margin=margin(r=10, b=-6, unit="pt")),
  autoplot(ts(season.168$trend, frequency=168)) + labs(title="Trend", cex=0.7) + ylab("") + xlab("") + theme(plot.margin=margin(l=4, b=-6, unit="pt")),
  autoplot(ts(season.168$figure, frequency=1)) + labs(title="Weekly seasonality", cex=0.7) + ylab("") + xlab("") + theme(plot.margin=margin(r=10, b=-6, unit="pt")),
  autoplot(aggregated-season.24$seasonal-season.168$seasonal-season.168$trend) + labs(title="Residuals", cex=0.7) + ylab("") + xlab("") + theme(plot.margin=margin(t=-6, b=-6, unit="pt")),
  nrow = 3,
  greedy = T)

@@

plot_grid(
  autoplot(ts(seasonbase.24$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct daily"), size=1.5) +
    geom_line(aes(y=season.24$figure, col="Multi daily")) +
    scale_colour_manual("", values=c("purple", "yellow"), breaks=c("Direct daily", "Multi daily")),
  autoplot(ts(seasonbase.168$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct weekly"), size=1.5) +
    geom_line(aes(y=season.168$figure + rep(season.24$figure, 7), col="Multi w+d")) +
    #geom_line(aes(y=rep(seasonbase.24$figure, 7), col="Direct daily"), size=1.5) +
    #geom_line(aes(y=rep(season.24$figure, 7), col="Multi daily")) +
    geom_line(aes(y=season.168$figure, col="Multi weekly")) +
    scale_colour_manual("", values=c("red", "purple", "blue", "yellow", "green"), breaks=c("Direct weekly", "Direct daily", "Multi" %s+% c(" weekly", " daily", " w+d"))),
  nrow = 2)

@@

```{r diffs acf/pacf}
seis.h.train %>% ggtsdisplay(lag.max = 168*5)
seis.h.train %>% diff() %>% ggtsdisplay(lag.max = 168*5)
# seis.h.train %>% diff(D=2) %>% ggtsdisplay(lag.max = 168)
# seis.h.train %>% diff(lag=24) %>% ggtsdisplay(lag.max = 168)
# seis.h.train %>% diff() %>% diff(lag=24) %>% ggtsdisplay(lag.max = 24, main="Supposed model: ARIMA(24,1,0)(0,1,2)[24]") # ARIMA(0,1,0)(0,1,2)[24]
# previous => ARIMA(5,1,0)(0,1,7)[24]
# seis.h.train %>% diff() %>% diff(lag=24, D=2) %>% ggtsdisplay(lag.max = 168)
seis.h.train %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5) # arima (0,0,0)(0,1,1/2)[168]
# acf variable and partially sinusoidal, pacf shows peaks each 168
seis.h.train %>% diff() %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5) # arima (0,1,0)(0,1,1)[168]
# acf cuts off after p=168, pacf shows peaks each 168
# seis.h.train %>% diff() %>% diff(lag=168, D=2) %>% ggtsdisplay(lag.max = 168)
seis.h.train %>% diff() %>% diff(lag=24) %>% diff(lag=168) %>% ggtsdisplay(lag.max = 168*5)
```

```{r plot_forecast func}
plot_forecast <- function(forec) {
  BP <- Box.test(forec$model$residuals, type="Box-Pierce")
  autoplot(forec) +
    labs(title=forec$method, subtitle = paste0("Box-Pierce p-value=", print_dec(BP$p.value, decimals=4))) +
    autolayer(forec$mean, series="Forecast") -> plotted
  return(plotted)
}
```

```{r arima}
to_test <- list(
  #c(0,0,0,0,0,1),  # BAD, Box-Pierce = 0
  #c(0,0,0,0,0,2),  # BAD, Box-Pierce = 0
  #c(0,0,0,0,1,1),  # nice, Box-Pierce = 0
  c(0,0,0,0,1,2),
  #c(0,0,0,0,1,3),  # CAN'T BE DONE: 3*168 = 504 > 350 (limit of arima on r)
  #c(0,1,0,0,0,1),  # BAD, Box-Pierce = 0
  c(0,1,0,0,1,1),
  c(0,1,0,0,1,2),
  c(2,0,0,0,1,0),  # hint from auto.arima D=1
  c(5,0,0,0,1,0),  # hint from auto.arima D=1
  c(3,1,1,0,1,0),  # hint from auto.arima D=1 d=1
  c(3,1,1,0,1,1),  # after hint then ACF + PACF
  c(4,1,2,0,1,0),  # hint from auto.arima d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10
  c(4,1,2,0,1,1)  # after hint then ACF + PACF
)
if (!exists("models")){
  models <- list()
}
for (model in to_test) {
  # model <- c(4,1,2,0,1,0)
  fit.name <- paste0("fit", paste0(model, collapse = ""))
  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.train) %s+% "]")
  if (!(fit.name %in% names(models))) {
    models[[fit.name]] <- Arima(seis.h.train, model[1:3], model[4:6])
  }
  if (!(forecast.name %in% names(models))) {
    models[[forecast.name]] <- forecast(models[[fit.name]], h=168*18)
  }
}
#models[["fitauto"]] <- auto.arima(seis.h.train, D=1, approx=F, trace = T)
#models[["forecastauto"]] <- forecast(models[["fitauto"]], h=168*18)  
#models[["fit311010"]] <- models[["fitauto"]]
#models[["fit412010"]] <- fit
#models[["forecast412010"]] <- forecast(models[["fit412010"]], h=168*18)

@@

checkresiduals(plotting.model)
residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
plot_grid(residuals.acf, residuals.pacf, nrow=2)
plot_forecast(plotting.forecast) + 
  autolayer(seis.h.test, series="Original")
plots <- show_overlap_and_diff(
  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
  base.desc = "original series",
  compare.desc = "reconstructed series",
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="% over ts span",
  diff.ylab.sec.accuracy = .01
)
plot_grid(plots[[1]], plots[[2]], nrow = 2)
accuracy(plotting.forecast, seis.h.test)

plotting.model <- models[["fit010012"]]
plotting.forecast <- models[["forecast010012"]]
checkresiduals(plotting.model)
residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
plot_grid(residuals.acf, residuals.pacf, nrow=2)
plot_forecast(plotting.forecast) + 
  autolayer(seis.h.test, series="Original")
plots <- show_overlap_and_diff(
  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
  base.desc = "original series",
  compare.desc = "reconstructed series",
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="% over ts span",
  diff.ylab.sec.accuracy = .01
)
plot_grid(plots[[1]], plots[[2]], nrow = 2)
accuracy(plotting.forecast, seis.h.test)

plotting.model <- models[["fit412010"]]
plotting.forecast <- models[["forecast412010"]]
checkresiduals(plotting.model)
residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
plot_grid(residuals.acf, residuals.pacf, nrow=2)
plot_forecast(plotting.forecast) + 
  autolayer(seis.h.test, series="Original") 
plots <- show_overlap_and_diff(
  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
  base.desc = "original series",
  compare.desc = "reconstructed series",
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="% over ts span",
  diff.ylab.sec.accuracy = .01
)
plot_grid(plots[[1]], plots[[2]], nrow = 2)
accuracy(plotting.forecast, seis.h.test)

plotting.model <- models[["fit500010"]]
plotting.forecast <- models[["forecast500010"]]
checkresiduals(plotting.model)
residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
plot_grid(residuals.acf, residuals.pacf, nrow=2)
plot_forecast(plotting.forecast) + 
  autolayer(seis.h.test, series="Original") 
plots <- show_overlap_and_diff(
  seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
  base.desc = "original series",
  compare.desc = "reconstructed series",
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="% over ts span",
  diff.ylab.sec.accuracy = .01
)
plot_grid(plots[[1]], plots[[2]], nrow = 2)
accuracy(plotting.forecast, seis.h.test)

for (model in to_test) {
  # model <- c(4,1,2,0,1,0)
  fit.name <- paste0("fit", paste0(model, collapse = ""))
  forecast.name <- paste0("forecast", paste0(model, collapse = ""))
  print("SARIMA(" %s+% paste0(model[1:3], collapse = ",") %s+% ")(" %s+% paste0(model[4:6], collapse = ",") %s+% ")[" %s+% frequency(seis.h.train) %s+% "]")
  if ((fit.name %in% names(models))) {  if ((forecast.name %in% names(models))) {
    plotting.model <- models[[fit.name]]
    plotting.forecast <- models[[forecast.name]]
    checkresiduals(plotting.model)
    residuals.acf <- autoplot(acf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
    residuals.pacf <- autoplot(pacf(plotting.model$residuals, na.action = na.pass, plot=F, lag.max=168*5)) + scale_x_continuous() + labs(title = paste0("Series: residuals of ", plotting.forecast$method))
    plot_grid(residuals.acf, residuals.pacf, nrow=2)
    plot_forecast(plotting.forecast) + 
      autolayer(seis.h.test, series="Original") 
    plots <- show_overlap_and_diff(
      seis.h.train, plotting.forecast$fitted, xlab=seasonality.xlab,
      base.desc = "original series",
      compare.desc = paste0("refit with ", plotting.forecast$method),
      overlap.ylab = "Mean hourly movement (nm)",
      diff.title = "Difference between ts",
      diff.ylab = "nm", diff.ylab.sec="% over ts span",
      diff.ylab.sec.accuracy = .01
    )
    plot_grid(plots[[1]], plots[[2]], nrow = 2)
    accuracy(plotting.forecast, seis.h.test)
  }}
}
qqPlot(models[["fit500010"]]$residuals)
```

```{r}
#fit <- auto.arima(seis.h.train, d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10, trace = T, parallel = T)
```
# D=1 m=24
 Fitting models using approximations to speed things up...

@@

# d=1, D=1, max.p=5, max.q=2, max.P=5, max.Q=2, max.order=10 m=168
 Fitting models using approximations to speed things up...
